---
title: "131 Homework 1"
author: "Tonia Wu"
date: "3/29/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
#install.packages('forcats')
#install.packages('corrplot')
library(forcats)
library(corrplot)
```

**Q1**. *Define supervised and unsupervised learning. What are the difference(s) between them?*

- Supervised learning involves modelling to predict an output based on input(s). This requires training data from which the model will learn.
- Unsupervised learning involves inputs and no outputs so that we can learn about the data and potentially discover patterns

**Q2**. *Explain the difference between a regression model and a classification model, specifically in the context of machine learning.*

- Regression models involve continuous, quantitative output
- Classification models involve qualitative output (such as a yes/no resu)

**Q3**. *Name two commonly used metrics for regression ML problems. Name two commonly used metrics for classification ML problems.*
- Regression: 
- Classification: whether something will increase or decrease
!!! doesn't look right

**Q4**. *As discussed, statistical models can be used for different purposes. These purposes can generally be classified into the following three categories. Provide a brief description of each.*

- *Descriptive models*:

- *Inferential models*:

- *Predictive models*:

**Q5**. *Predictive models are frequently used in machine learning, and they can usually be described as either mechanistic or empirically-driven. Answer the following questions.*

|     *Define mechanistic. Define empirically-driven. How do these model types differ? How are they similar?*

|           - A mechanistic model takes a relationship  or theory and imposes it on the data (from lecture)
|           - AAn empirically-driven model looks at the data and sees what best fits it (from lecture)
|           - ASimilarities: !!!!!!!!!!!!!!!
|           - ADifferences: !!!!!!!!!!!!!!


|     *In general, is a mechanistic or empirically-driven model easier to understand? Explain your choice.*


|     *Describe how the bias-variance tradeoff is related to the use of mechanistic or empirically-driven models.*


**Q6**. *A political candidate’s campaign has collected some detailed voter history data from their constituents. The campaign is interested in two questions:*

|     a. *Given a voter’s profile/data, how likely is it that they will vote in favor of the candidate?*

|     b. *How would a voter’s likelihood of support for the candidate change if they had personal contact with the candidate?*

|     *Classify each question as either predictive or inferential. Explain your reasoning for each.*
|           - a: as we are using past data to predict the future, this involves a predictive model.
|           - b: since we're interested in the relationship between the inputs and the output, this involves an 
|                inferential model.

**Exploratory Data Analysis: mpg**

**E1**. *We are interested in highway miles per gallon, or the hwy variable. Create a histogram of this variable. Describe what you see/learn.*
|       Looks like we're dealing with nonsymmetric bimodal data:

```{r}
ggplot(data = mpg, aes(hwy)) + geom_histogram()
```

**E2**. *Create a scatterplot. Put hwy on the x-axis and cty on the y-axis. Describe what you notice. Is there a relationship between hwy and cty? What does this mean?*

There is an incredibly strong positive relationship between hwy (highway mpg) and cty (city mpg). Thus, as hwy increases, cty increases.

```{r}
ggplot(mpg, aes(x = hwy, y = cty)) + geom_point()
``` 

**E3**. *Make a bar plot of manufacturer. Flip it so that the manufacturers are on the y-axis. Order the bars by height. Which manufacturer produced the most cars? Which produced the least?*


```{r}
ggplot(data = mpg, aes(x = fct_infreq( manufacturer))) + 
  geom_bar(stat = 'count') +
  coord_flip()
```

**E4**. *Make a box plot of hwy, grouped by cyl. Do you see a pattern? If so, what?*

Looks like highway mpg decreases as the number of cylinders increases.

```{r}
hwy <- mpg$hwy
cyl <- mpg$cyl
boxplot(hwy ~ cyl)
```


**E5**. *Use the corrplot package to make a lower triangle correlation matrix of the mpg dataset.*

*Which variables are positively or negatively correlated with which others? Do these relationships make sense to you? Are there any that surprise you?*

Strong positive: displacement & cylinders, highway mpg & city mpg

Neither of these are surprising since I'd expect highway and city mpg to be quite similar, and engine displacement depends on cylinders in the first place.

Strong negative: highway mpg & displacement, city mpg & displacement, city mpg & cylinder, highway mpg & cylinder

For these, I tried Googling what effect cylinders have on gas usage, but came away with no clear answer. So I am surprised by this result, but mainly because I don't understand cars.

Weak positive correlation: year & displacement, cylinder & year

Little to no correlation: city & year, highway mpg & year

The lack of correlation between the mpgs and year was surprising to me because I thought mpg would increase over time. 

```{r}
# select only numeric variables
# source: statistics globe
mpg_2 <- select_if(mpg, is.numeric)
mpg_2
M <- cor(mpg_2)
corrplot(M, method = 'number', type = 'lower')
```










